---
title: "Ch3-dataVisualization"
author: "julia"
date: "2022/7/26"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)

```


`mpg` contains observations collected by the US Environmental Protection Agency on 38 models of car.
```{r}
ggplot2::mpg
view(mpg)
```

`manufacturer`: 製造商名稱
`model`: 模型名稱
`displ` : 發動機排量，升
`year`: 製造年份
`hwy`: 氣缸數
`trans`: 傳輸類型
`drv`: the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
`cty`: 每加侖城市英里數
`hwy`: 每加侖公路英里數
`fl` : 汽油種類
`class`: 車型

##3.2.2 Creating a ggplot
```{r}
ggplot(data=mpg)+
  geom_point(
    mapping=aes(
      x=displ,
      y=hwy
    )
  )
```
The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy).


##3.2.3 A graphing template
Basic Template: 
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))


##3.2.4 Exercises
1. Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data=mpg)
```
2. How many rows are in mpg? How many columns?
```{r}
?mpg
#A data frame with 234 rows and 11 variables
```

3. What does the drv variable describe? Read the help for ?mpg to find out.
`drv`: the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4. Make a scatterplot of hwy vs cyl.
```{r}
ggplot(data=mpg)+
  geom_point(
    mapping=aes(
      x=cyl,
      y=hwy
    )
  )
```

5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(data=mpg)+
  geom_point(
    mapping=aes(
      x=class,
      y=drv
    )
  )
```

#3.3 Aesthetic mappings

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=hwy,
    color=class
  ))
```

##3.3.1 Exercise
1. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```
因為這是透過手動，不包含變數的變動，就要放在aesthetic之外。

2. Which variables in `mpg`are categorical? Which variables are continuous? (Hint: type`?mpg`to read the documentation for the dataset). How can you see this information when you run`mpg`?
```{r}
names(mpg)
class(mpg$manufacturer)
class(mpg$displ)
```
3. Map a continuous variable to`color`,`size`, and`shape`. How do these aesthetics behave differently for categorical vs.continuous variables?
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(
    x=manufacturer,
    y=displ,
    color=model
  ))
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(
    x=manufacturer,
    y=displ,
    color=cty
  ))
```
4. What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=displ,
    color=displ
  ))
```
5. What does the`stroke`aesthetic do? What shapes does it work with? (Hint: use`?geom_point`)
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=hwy
  ),
  color="blue",
  size=3,
  stroke=1)
```
6. What happens if you map an aesthetic to something other than a variable name, like`aes(colour = displ < 5)`? Note, you’ll also need to specify x and y.
```{r}
ggplot(mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=hwy,
    color=displ<5
  ))
```

#3.5 Facets
刻面：可以呈現多個不同的圖
```{r}
ggplot(mpg)+
  geom_point(mapping=aes(
  x=displ,
  y=hwy
  ))+
  facet_wrap(~manufacturer)
```
```{r}
ggplot(mpg)+
  geom_point(mapping=aes(
  x=displ,
  y=hwy
  ))+
  facet_wrap(drv~cyl)
```

##3.5.1 Exercise
1. What happens if you facet on a continuous variable?
```{r}
ggplot(mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=cty
  ))+
  facet_wrap(~hwy)
```

2. What do the empty cells in plot with `facet_grid(drv ~ cyl)`mean? How do they relate to this plot?
```{r}
ggplot(mpg)+
  geom_point(mapping=aes(
    x=displ,
    y=hwy
  ))+
  facet_grid(drv~cyl)
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl))
```

3. What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)


ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
簡單來說就是.的那邊沒東西，看你要用值得還是橫的

4. Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

可以更清晰看到不同面相下的差異，但就比較無法進行比較。

5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

ans. /Number of rows and columns.


6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
因為通常圖片是橫的，用欄可以裝比較多。


#3.6 Geometric objects 幾何物件

```{r}
ggplot(data = mpg, mapping=aes(x=displ, y=hwy)) +
  geom_smooth()+
  geom_point(mapping=aes(color=class))
```
```{r}
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)
```
##3.6.1 Exercise
1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
`line chart`
```{r}
ggplot(mpg, mapping=aes(x=displ,y=hwy))+
  geom_line()
```
`boxplot chart`
```{r}
ggplot(mpg, mapping=aes(x=class,y=hwy))+
  geom_boxplot()
```
`histogram chart`
```{r}
ggplot(mpg, mapping=aes(x=displ,y=hwy))+
  geom_col()
```
`area chart`
```{r}
ggplot(mpg, mapping=aes(x=displ,y=hwy))+
  geom_area()
```

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
應該就是把圖例弄不見
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
4. What does the se argument to geom_smooth() do?
se: standard error
原本geom_smooth會有一層淡淡的are在下面，加上se後就能把他去掉
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy),
    se=FALSE)
```
5. Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
都一樣，寫在前面可以全部通用。

6. Recreate the R code necessary to generate the following graphs.
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth(se=FALSE)

ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth(mapping=aes(group=drv),se=FALSE)
```
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy,color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)

ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=drv))+
  geom_smooth(se=FALSE)
```
```{r}
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=drv))+
  geom_smooth(mapping=aes(linetype=drv),se=FALSE)
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),color="gray",size=2.8)+
  geom_point(mapping=aes(x=displ,y=hwy,color=drv))
```

#3.7 Statistical Transformations 統計轉換
```{r}
View(diamonds)
```
About this data set:
contains information about ~54,000 diamonds, including the `price`, `carat`, `color`, `clarity`, and `cut` of each diamond. 
```{r}
ggplot(diamonds)+
  geom_bar(
    mapping=aes(x=cut)
  )
```
1. 覆蓋默認的stat
```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(demo)+
  geom_bar(mapping = aes(x=cut))

ggplot(demo)+
  geom_bar(mapping = aes(x=cut,y=freq),stat="identity")
```
```{r}
ggplot(diamonds)+
  geom_bar(mapping = aes(x=cut,y=stat(prop),group=1))
```
```{r}
ggplot(diamonds)+
  stat_summary(
    mapping = aes(x=cut,y=depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

##3.7.1 Exercise
1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

2. What does geom_col() do? How is it different to geom_bar()?
col的default stat 是 identity；而bar的是count。
```{r}
ggplot(diamonds)+
  geom_col(mapping=aes(x=cut,y=depth))
```
3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
`geom_bar()`:stat_count()
`geom_boxplot()`:stat_boxplot()
`geom_smooth()`:stat_smooth()
`geom_density()`:stat_density()
`geom_bin_2d()`:stat_bin_2d()
`geom_contour()`:stat_contour()
`geom_count()`:stat_sum()
`geom_function()`:stat_function()
`geom_qq_line()`:stat_qq_line()
其實名字幾乎都一樣

4. What variables does stat_smooth() compute? What parameters control its behaviour?


5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?
因為geom_bar的default是用x項分group，而y軸用%時，也會一起分項計算，所以要將group統一一起計算才會有我們想要的。
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop),group=1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```

#3.8 Position adjustments

```{r}
ggplot(diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity))
```
我們會發現原本單純用scatterplot時有一些點會重疊，這時候我們可以用jitter使點稍微分開一點。
`position = "jitter"` adds a small amount of random noise to each point. This spreads the points out because no two points are likely to receive the same amount of random noise.
```{r}
ggplot(data=mpg)+
  geom_point(
    mapping=aes(
      x=displ,
      y=hwy
    ),
    position="jitter"
  )

```
##3.8.1 Exercise

1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
#這就是前面提到的問題，因為過度重複，難以有效辨別數據的關聯
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position="jitter")
#稍微用jitter使其小規模不準確，但可更有效呈現
```
2. What parameters to `geom_jitter()` control the amount of jittering?
透過width和height調整。
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter(
    width = 1,
    height= 1
  )
```

3. Compare and contrast `geom_jitter()` with `geom_count()`.
後者用大小表示多寡。
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```
4. What’s the default position adjustment for `geom_boxplot()`? Create a visualisation of the mpg dataset that demonstrates it.
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot(position="dodge")
```


























