---
title: 'Ch5 Data Transformation'
author: "julia"
date: "2022/10/17"
output: html_document
---

使用dplyr進行資料變換
本章會運用到來自nycflights13資料及進行dplyr練習
```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
view(nycflights13::flights)
flights
```
*?flights*
columns:
`year, month, day`
`Date of departure.`
`dep_time, arr_time`
`sched_dep_time, sched_arr_time`
`dep_delay, arr_delay`: in minutes
`carrier`
`flight`: Flight number.
`tailnum`: Plane tail number.
`origin, dest`: Origin and destination.
`air_time`: Amount of time spent in the air, in minutes.
`distance`: Distance between airports, in miles.
`hour, minute`: Time of scheduled departure broken into hour and minutes.
`time_hour`: Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.

*每個欄位名稱下面都有三或四個字母縮寫，他們描述每個變數的類型：*
- int - 整數integers
- dbl - doubles雙精度浮點數或實數real numbers
- chr - 字元向量character vectors或字串strings
- dttm - 日期與時間
- lgl - 邏輯向量logical vectors，只有true/false
- fctr - 因子factors，R以固定幾個可能值來表示類別變數(categorical variables)的一種資料結構

#5.1 dplyr Basics

*dplyr五個基礎函式*
- filter(): 依據觀察的值進行挑選(列)
- arrange(): 重新安排資料列的順序
- select(): 以名稱挑選變數(欄)
- mutate(): 以既有變數的函式運算值建立新的變數
- summarize(): 從許多值產生單一的摘要資訊
這些函式的共通點是動詞的運作方式：
1. 第一個引數(argument)是一個資料框(data frame)
2. 後面的引述用不帶引號的變數名稱描述要對資料框做的事情
3. 結果會是一個新的資料框

group_by()可以搭配以上使用。

#5.2 filter()
*filter(data, column1 == 1, column2 == 2)*

```{r}
Sep23 <- filter(flights, month==9, day==23)
Sep23
```
##Boolean布林運算
這邊蠻必要的，之後邊看邊練習。
##缺失值NA
check if it's NA:
```{R}
is.na(flights$year)
#都是false,表示都非na
```
filter只會留下條件符合的值，其他不符合的(false)和NA都會被刪掉，所以要留下NA要明確要求：
```{R}
df <- tibble(x=c(1,NA,3))
filter(df,x>1)
filter(df,is.na(x)|x>1)
```
##Exercise
1. 找出符合下列條件的所有班機:
a. 抵達誤點超過兩小時
```{r}
filter(flights, arr_delay>=120)
```
b. 飛往houston(iah或hou)
```{r}
filter(flights, dest=="IAH"|dest=="HOU")
filter(flights, dest%in%c("IAH","HOU"))
```
c. 屬於United,American或Delta航空公司
```{R}
filter(flights, carrier%in%c("UA","AA","DL"))
```
d. 在夏天(7-9月)出發
```{r}
filter(flights,month%in%c(7,8,9))
filter(flights,month%in%7:9)
```

e. 晚了兩小時以上抵達但沒有晚出發
```{r}
filter(flights, arr_delay>=120&dep_delay<=0)
filter(flights, arr_delay>=120, dep_delay<=0)
```
f. 誤點至少一小時，但在飛行時以補超過30分鐘時間
```{r}
filter(flights, dep_delay>=60, dep_delay-arr_delay>=30)
```
g. 在子夜(midnight)和6am之間出發
```{r}
#filter(flights, dep_time>=0, dep_time<=600) ##錯誤
filter(flights, dep_time==2400|dep_time<=600)
```
**In dep_time, midnight is represented by 2400, not 0. You can verify this by checking the minimum and maximum of dep_time.

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
filter(flights,between(month,7,9))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights,is.na(dep_time))
#dep_delay, arr_time, arr_delay, air_time are missing
#班機取消?
```
4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}
NA^0  #all numeric values  X^0=1
NA|TRUE   #because "anything or TRUE" is TRUE
FALSE&NA   #NA & FALSE is FALSE because "anything and FALSE" is always FALSE
NA*0    #Since x∗0=0 for all finite numbers we might expect NA * 0 == 0, but that’s not the case. The reason that NA * 0 != 0 is that 0*∞ and 0*−∞ are undefined. R represents undefined results as NaN, which is an abbreviation of “not a number”.
```















