# Categorical data

[***總覽課程內容：***]{.ul} group使用、data的wide和long form、 factor用來排列資料在plot上的順序、 position的使用(stack)、

## Aesthetics: group

```{r}
ggplot() +
  geom_line(
    mapping=aes(
      x=c(1, 2, 3),
      y=c(2, 3, 2),
    )
  ) +
  geom_line(
    mapping=aes(
      x=c(1, 2, 3),
      y=c(5, 2, 6)
    )
  )

```

Use `group` aesthetic to combine

-   multiple same geom layers

into one.

```{r}
ggplot() +
  geom_line(
    mapping=aes(
      x=c(1, 2, 3, 1, 2, 3),
      y=c(2, 3, 2, 5, 2, 6),
      group=c("m", "m", "m", "f", "f", "f")
    )
  )

```

```{r}
ggplot() +
  geom_line(
    mapping=aes(
      x=c(1, 2, 3, 1, 2, 3),
      y=c(2, 3, 2, 5, 2, 6),
      group=c("m", "m", "m", "f", "f", "f"),
      color=c("m", "m", "m", "f", "f", "f")
    )
  )

```

-   Any aesthetic differentiates group can replace group.若顏色跟group的分群一樣，可以省略group。

```{r}
ggplot() +
  geom_line(
    mapping=aes(
      x=c(1, 2, 3, 1, 2, 3),
      y=c(2, 3, 2, 5, 2, 6),
      # group=c("m", "m", "m", "f", "f", "f"),
      color=c("m", "m", "m", "f", "f", "f")
    )
  )

```

::: {.alert .alert-warning}
-   When there is no aesthetic mapping to differentiate groups, use `group` aesthetic mapping.
:::

## Geom overlapping

When geom layers overlap, we can use

-   `alpha` aesthetic.

If multiple geometries are created within the one `geom_` call (using grouping aesthetics), we can also set

-   position: "stack", "dodge" or "jitter" (some of them might not apply to certain `geom_`)

```{r , eval=T, echo=F}
teachDS::img_centering("https://clauswilke.com/dataviz/visualizing_proportions_files/figure-html/health-vs-age-1.png")
```

-   <https://clauswilke.com/dataviz/visualizing-proportions.html#fig:health-vs-age>

-   x is continuous, or discrete with many types

-   y the cumulative proportion

------------------------------------------------------------------------

```{r}
ggplot() +
  geom_area(
    mapping=aes(
      x=c(1, 2, 3),
      y=c(0.2, 0.3, 0.2),
    )
  ) +
  geom_area(
    mapping=aes(
      x=c(1, 2, 3),
      y=c(0.4, 0.3, 0.52) + c(0.2, 0.3, 0.2) # the additive is for accumulative purpose
    ), 
    alpha=0.5 #the former layers will be invisible if alpha=1
  )
```

## Position: stack

-   我們常會遇到的問題可能是資料們之間會覆蓋put y on top of the overlapping geom's y

-   create accumulative result. 所以position有兩種選項："identity","stack"，後者會將每個資料往上堆疊。

```{r}
ggplot() +
  geom_area(
    mapping=aes(
      x=c(1, 2, 3, 
        1, 2, 3),
      y=c(0.2, 0.3, 0.2, 
        0.4, 0.3, 0.52),
      fill=c("m", "m", "m", 
        "f", "f", "f")
    ),
    position="stack" #input$position
  )
```

-   stack position is accumulative; no need to compute the accumulative value yourself.

-   the default position in `geom_area` is "stack". Therefore, you can omit position argument.

------------------------------------------------------------------------

```{r data_cat1}
data_cat1 <- data.frame(
      x=c(1, 2, 3, 1, 2, 3),
      y=c(0.2, 0.3, 0.2, 0.4, 0.4, 0.52),
      fill=c("m", "m", "m", "f", "f", "f")
)
```

```{r}
ggplot(
  data=data_cat1
) + 
  geom_area(
    mapping=aes(
      x=x,
      y=y,
      fill=fill
    )
  )
```

::: {.alert .alert-warning}
When aesthetic mapping involves with unordered data, it will

-   convert the data series into factor (unless the series is already a factor);

-   conduct the mapping according to the level sequence of the converted factor.
:::

```{r}
#|>的用法：放進後面函式的()裡面
class(data_cat1) #和後面的功能一樣
data_cat1$fill |>
  factor() |>
  levels() |>
  class()
```

雖然是character，本身沒有排序，但r在判定時會給先後順序

```{r}
data_cat1$fill |>
  factor() -> fill_asFactor #先把他轉成factor看其順序
fill_asFactor
levels(fill_asFactor) #levels可以顯示，但不會告訴你順序
```

## Factor

When grouping aesthetics vary the look of geometries across different groups of data, it is crucial that users declare the mapped series with proper class.

-   `factor(data_series, levels)` parses `data_series` into a categorical data with expressing sequence defined by `levels`. 順序可以透過factor去改變。

-   If omit `levels` the level sequence will be determined by the collateral sequence defined by your operating system.

```{r}
ggplot(
  data=data_cat1
) + 
  geom_area(
    mapping=aes(
      x=x,
      y=y,
      group = factor(fill, levels = c("m", "f")),
      fill=factor(fill, levels=c("m", "f"))
    ) #fill是顏色的順序先後有調整而已，若想要調整資料上下的順序還是要靠group
  )
```

-   Here we declare factor on-the-go.

We can also declare factor in the data frame first也可以在data裡就先處理好:

```{r}
data_cat1_copy <- data_cat1
data_cat1_copy$fill |>
  factor(levels=c("m", "f")) -> 
  data_cat1_copy$fill
```

-   `|>` is a R 4.0+ equipped operator, which makes:

```{r}
f(x, ....) # equivalent to
x |> f(...)
```

```{r}
ggplot(
  data=data_cat1_copy
) + 
  geom_area(
    mapping=aes(
      x=x,
      y=y,
      fill=fill
    ),
    #position="stack"
  )
```

###以後只要data中有要分類，有字串的話，都依定要用factor，這樣順序才不會到不同電腦中都亂跑。

## Proportional data

```{r}
data_cat2_wide <- data.frame(
      x=c(1, 2, 3), #不同年齡層
      y_a=c(0.2, 0.3, 0.2), 
      y_b=c(0.4, 0.4, 0.52),
      y_c=c(0.4, 0.3, 0.28) #不同身體健康狀態
)

data.frame(
  x=c(1,2,3,1,2,3,1,2,3),
  values_to =c(0.2, 0.3, 0.2, 0.4, 0.4, 0.52, 0.4, 0.3, 0.28),
  fill = c("y_a","y_a","y_a","y_b","y_b","y_b","y_c","y_c","y_c")
)

#如果要用到美學mapping要group的話，data frame一定要是用long form
#通常像要wide在計算不同類別加總時比較好計算；而long的話就是在跑計量或統計時比較好寫，因為較不繁瑣。
data_cat2_wide |> 
  tidyr::pivot_longer(
    cols=y_a:y_c,
    names_to = "fill",
    values_to= "y"
  ) ->
  data_cat2

View(data_cat2)
```

```{r}
ggplot(
  data=data_cat2
) + 
  geom_area(
    mapping=aes(
      x=x,
      y=y,
      group=factor(fill, levels=c("y_c","y_a","y_b")),
      fill=fill
    ),
    color="white" #邊線顏色
  )
```

------------------------------------------------------------------------

When x mapping series has limited cases and is discrete, a bar chart with position dodge is better. #dodge要左右散開不重疊，都從0開始畫

```{r}
ggplot(
  data=data_cat2
) + 
  geom_col(
    mapping=aes(
      x=x,
      y=y,
      group = factor(fill, levels = c("y_a", "y_b", "y_c")),
      fill= factor(fill, levels=c("y_a", "y_b", "y_c")
    ),
    color="white",
    width=0.8, #input$width
    size=1, #input$size
    position = "dodge" #input$position
  ) 
```

-   `width`: the width of the bar

-   `size`: the size of the stroke

------------------------------------------------------------------------

Pie chart:

-   not good for comparing proportion across more than one dimension

```{r}
library(dplyr)
data_cat2 %>%
  filter(
    x==1
  ) -> 
  data_cat2_x1only
```

```{r}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  )
```

```{r piechart}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  coord_polar(
    theta = "y"
  )
```

## Adding text

adding text

```{r}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  geom_text(
    aes(
      x=x,
      y=y,
      label=fill
    ),
    position = "stack"
  )
```

-   `geom_col` stack sequence is based on `fill` level sequence.

-   `geom_text` stack sequence is based on observation sequence.

::: {.alert .alert-info}
Grouping aesthetics determine the sequence of stacking. In `geom_col`, `fill` is the grouping aesthetic. To make `geom_text` stack labels in sequence as `fill` in `geom_col`, we can put `group=fill` in `geom_text` to create such a sequence.

```{r}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  geom_text(
    aes(
      x=x,
      y=y,
      label=fill,
      group=fill
    ),
    position = "stack"
  )
```
:::

Change labels to represent the proportion values of `y`

```{r}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  geom_text(
    aes(
      x=x,
      y=y,
      label=y, # use y to label now
      group=fill
    ),
    position = "stack"
  )
```

-   `position` argument also takes position functions.

-   When you know what type of position you want, you can use corresponding position function to fine tune the position.

```{r}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  geom_text(
    aes(
      x=x,
      y=y,
      label=y,
      group=fill
    ),
    position = position_stack(vjust=0.5)
  )
```

```{r piechart2}
ggplot(
  data=data_cat2_x1only
) + 
  geom_col(
    aes(
      x=x,
      y=y,
      fill=fill
    )
  ) +
  geom_text(
    aes(
      x=x,
      y=y,
      label=y,
      group=fill
    ),
    position = position_stack(vjust=0.5)
  ) +
  coord_polar(
    theta = "y"
  ) +
  theme_void()
```

------------------------------------------------------------------------

When x-axis is also representing a categorical data:

```{r dd}
dy=0.03 # input$dy
ggplot(
  data=data_cat2
) + 
  geom_col(
    mapping=aes(
      x=x,
      y=y,
      fill=fill
    ),
    color="white",
    width=0.8, #input$width
    position = "dodge" #input$position
  )+
  geom_text(
    mapping=aes(
      x=x,
      y=y-dy,
      group=fill,
      label=y
    ),
    size=8, #input$size
    position=position_dodge(width=
        0.8 #input$dodge
        )
  )
```

-   text position_dodge has the same width as `geom_col` to ensure the same dodging distance.

## More on position

<https://ggplot2.tidyverse.org/reference/index.html#section-position-adjustment>

## Coordination flip

```{r, eval=T, echo=F}
teachDS::img_centering("https://www.economist.com/img/b/1280/759/90/sites/default/files/20191130_WOC194.png")
```

```{r}
ggplot()+
  geom_col(
    mapping=
      aes(
        x=c("A", "B", "C"),
        y=c(56, 77, 92)
      )
  )+
  coord_flip()
```

------------------------------------------------------------------------

Another common application of `coord_flip` is:

```{r , eval=T, echo=F}
teachDS::img_centering("img/us-presidential-election.jpg", width="100%")
```

```{r}
dx=4 #input$dx
h=0.5 #input$h
dt=0 #input$dt
ggplot()+
  geom_col(
    mapping=aes(
      x=c(1, 1),
      y=c(306, 232),
      fill=c("biden","trump")
    ),
    width=1
  )+
  geom_segment(
    mapping=aes(
      x=1-h,
      y=270,
      xend=1+h,
      yend=270
    )
  )+
  geom_text(
    mapping=aes(
      x=1+dt,
      y=270,
      label="270"
    ),
    size=8 #input$text
  )+
  xlim(1-dx, 1+dx)+ # make sure cover 0.5-1.5 so the bar width can be accomodate
  coord_flip()+
  theme_void()+
  theme(legend.position = "none")
```

## Summary

-   Grouping aesthetic separate a data frame into various subsample data frame and apply the `geom_` function to each one of them in the sequence determined by the mapping factor's levels sequence.

-   When `group` aesthetic and other aesthetic share the same mapping variable, `group` aesthetic can be ignored.

-   When deal with grouping variable, values of y from different groups at the same x can have position choice:

    -   "identity": respect ys as it is.\
    -   "stack": stack ys according to grouping level sequence.
    -   "dodge": respect ys as it is but move their x values left and right according to grouping level sequence.

## Exercise

### 1 {.unnumbered}

```{r, echo=F, eval=T}
teachDS::img_centering("https://images.squarespace-cdn.com/content/v1/55b6a6dce4b089e11621d3ed/1586289284165-6Z2M2XKD645EJRCV3TPV/Area+chart+example.png?format=1500w")
```

```{r}
set.seed(2020)
type1 <- function(){
  dnorm(1990:2010, mean=1991, sd=3)*2500 -> x
    round(x, digits = 0)
}
type2 <- function(){
  dnorm(1994:2010, mean=1998, sd=2)*1000->x
  round(x, digits=0)
}
type3 <- function(){
  dnorm(2002:2010, mean=2005, sd=2.3)*800->x
  round(x, digits=0)
}

data_set4 <- data.frame(
  year=c(
    1990:2010,
    1994:2010,
    2002:2010),
  storage_type=
    c(
      rep("type1", 21),
      rep("type2", 17),
      rep("type3", 9)),
  sales_amount=c(
    type1(), 
    type2(),
    type3()
  )
)
```

### 2 {.unnumbered}

```{r, echo=F, eval=T}
teachDS::img_centering("https://www.economist.com/img/b/1000/1339/90/sites/default/files/images/print-edition/20211023_FBC728.png")
```

### 3 {.unnumbered}

```{r, echo=F, eval=T}
teachDS::img_centering("https://www.economist.com/img/b/1000/2753/90/sites/default/files/20211023_WWC082.png")
```

### 4 {.unnumbered}

```{r, echo=F, eval=T}
teachDS::img_centering("http://uc.udn.com.tw/photo/2020/01/12/1/7324036.png")
```
